---
title: "Practicing Data Wrangling with Tree Equity Data"
author: "Jo Hardin -- AWM + GEM"
date: "October 30, 2015"
format:
  html:
    embed-resources: true
execute: 
  warning: false
  message: false
---

```{r}
library(tidyverse)
```


## Instructions

In this worksheet, you'll practice writing **ggplot R code** for creating data visualizations. 
Each question guides you through building layers of plots for the `ca_tes` dataset. 

**Don't stress yourself out over perfect syntax** -- the goal is to practice what you've learned about data wrangling so far.


# The Data

## Loading in the data

```{r}
ca_tes <- tempfile(fileext = ".zip") |>
  (\(f) { 
    download.file("https://tes-app-data-share.s3.amazonaws.com/ca/ca_csv.zip", f, quiet = TRUE)
    read_csv(unz(f, unzip(f, list = TRUE)$Name[1]))
  })()
```


## Looking at the data

```{r}
ca_tes
```

Variable descriptions are given (mostly) at <a href = "https://www.treeequityscore.org/methodology?tab=data-glossary" target = "_blank">https://www.treeequityscore.org/methodology?tab=data-glossary</a>


## An observation

> Observational Unit: object by which data is collected and analyzed

For example, an observational unit could be any of the following:

-   Person

-   County

-   State

-   Country

-   *What is the observational unit in the tree equity dataset?*



## What is the observational unit?

```{r}
ca_tes
```

-   The GEOID or census block group describes a particular community or neighborhood in a given area that contains between 600 - 3,000 people [^2]

[^2]: American Forests Methods and Data Glossary. https://www.treeequityscore.org/methodology?tab=data-glossary

## What are the variables?

-   `treecanopy`: actual tree coverage in an area
-   `tc_goal`: ideal tree coverage based on needs
-   `tc_gap`: how far behind the area is (`tc_goal` - `treecanopy`)
-   `pctpocnorm`: % residents who are people of color
-   `pctpovnorm`, `unemplnorm`: indicators of equity
-   `lingnorm`: % of non-English households (linguistic isolation)
-   `tes`: tree equity score (gap × need)



## 1. Tree Equity vs. Poverty

> Is there a relationship between poverty (`pctpov`) and tree equity score (`tes`)?


```{r}
#| echo: true
#| eval: false

ca_tes |>
  ___(___(x = ___, y = ___)) +
  ___(color = "navy", alpha = 0.05) 
```


```{r}
#| echo: true
#| eval: true
#| code-line-numbers: "1|2|3|4|5|6"

ca_tes |>
  ggplot(aes(x = pctpov, y = tes)) +
  geom_point(color = "navy", alpha = 0.05) + 
  labs(title = "Tree Equity vs. Poverty",
       x = "Percent in Poverty",
       y = "Tree Equity Score")
```

## 2. Distribution of tree equity scores

> What is the distribution of tree equity scores (`tes`)?


```{r}
#| echo: true
#| eval: false

___ |>
  ggplot(aes(___)) +
  ___(fill = "red")
```


```{r}
#| echo: true
#| eval: true
#| code-line-numbers: "3"

ca_tes |>
  ggplot(aes(x = tes)) +
  geom_histogram(fill = "red") +
  labs(title = "Distribution of tree equity scores",
       x = "Tree Equity Score")

```

## 3. Distribution of tree equity scores

> What is the distribution of tree equity scores (`tes`) created by a boxplot?


```{r}
#| echo: true
#| eval: false

___ |>
  ggplot(aes(___)) +
  ___(fill = "violet")
```



```{r}
#| echo: true
#| eval: true
#| code-line-numbers: "3"

ca_tes |>
  ggplot(aes(x = tes)) +
  geom_boxplot(fill = "violet") +
  labs(title = "Distribution of tree equity scores",
       x = "Tree Equity Score")

```



```{r}
ca_tes |> 
  ggplot(aes(x = tes, y = biome)) + 
  geom_boxplot()
```

## 4. Tree equity vs. linguistic isolation

> Is there a relationship between linguistic isolation (`lingnorm`) and `tes` in an area. 

  - Use ggplot to plot the relationship between `tes` and `lingnorm`
  - Choose the best plot for the variable types
  - Create fitting labels for your axes and plot title

##

```{r}
#| echo: true
#| eval: true
#| code-line-numbers: "1|2|3|4|5|6"

ca_tes |>
  ggplot(aes(x = lingnorm, y = tes)) +
  geom_point(color = "turquoise") + 
  labs(title = "Tree Equity vs. Linguistic Isolation",
       x = "Percent in Linguistic Isolation",
       y = "Tree Equity Score")
```



## 5. Wrangle + plot

Recreate the plot above, but summarize the information **per `county`** before plotting. That is, `Los Angeles County` should now only be represented by one dot. Summarize both `lingnorm` and `tes` using the `mean()` function.




```{r}
#| echo: true
#| eval: true
#| code-line-numbers: "1,2,3,4"

ca_tes |> 
  group_by(county) |> 
  summarize(ave_ling = mean(lingnorm),
            ave_tes = mean(tes)) |> 
  ggplot(aes(x = ave_ling, y = ave_tes)) + 
  geom_point() + 
  labs(title = "Tree Equity vs. Linguistic Isolation by County",
       x = "Percent in Linguistic Isolation",
       y = "Tree Equity Score")
```

## 6. Wrangle + plot + label

Recreate the plot above, but now add labels to the plot to indicate the county of each dot.  Use `geom_label()` or `geom_text()`. Try both! You'll need to use `label = county`...  do you need `aes()`? 




```{r}
#| echo: true
#| eval: true
#| code-line-numbers: "7"

ca_tes |> 
  group_by(county) |> 
  summarize(ave_ling = mean(lingnorm),
            ave_tes = mean(tes)) |> 
  ggplot(aes(x = ave_ling, y = ave_tes)) + 
  geom_point() + 
  geom_text(aes(label = county)) + 
  labs(title = "Tree Equity vs. Linguistic Isolation by County",
       x = "Percent in Linguistic Isolation",
       y = "Tree Equity Score")
```


## 7. Wrangle + plot + label

There is an R package called **ggrepel**.  It includes a geom called `geom_label_repel`. Try it!  (To use a function from a package, you can write: `ggrepel::geom_label_repel()`.)

**Hint:** Is there a yellow bar at the top asking if you want to load the R package **ggrepel**?  Say yes!




```{r}
#| echo: true
#| eval: true
#| code-line-numbers: "7"

ca_tes |> 
  group_by(county) |> 
  summarize(ave_ling = mean(lingnorm),
            ave_tes = mean(tes)) |> 
  ggplot(aes(x = ave_ling, y = ave_tes)) + 
  geom_point() + 
  ggrepel::geom_label_repel(aes(label = county)) + 
  labs(title = "Tree Equity vs. Linguistic Isolation by County",
       x = "Percent in Linguistic Isolation",
       y = "Tree Equity Score")
```

# Explore, Analyze, Present!


## Your turn 

> You will work in small groups to investigate tree equity in a new state(s) 

-   **Each group will:**   
  -   Choose a research question(s)    
  -   Use tidyverse wrangling tools to explore the data     
  -   Create plot(s) with **ggplot2**     

## Your goals  

-   Clearly state your research question  
-   Support your question with R code and graphs  
-   Use the data to identify a trend, pattern, or disparity  
-   Communicate **WHY** it matters 

## Sample questions

-   How does tree canopy vary across different neighborhoods in *chosen state*?  
-   Is there a relationship between percent in poverty and tree equity score in *chosen state*? 
-   How do tree equity trends compare between two or more *chosen states*?  
-   Are places with lower tree equity also more likely to have higher health disparities in *chosen state*?  
-   What if you break down / `summarize()` the information by county?

## Getting the data 

**Navigate to https://www.treeequityscore.org/:**

-   in drop down menu, click the "methods and data" tab 
-   choose the "data download" page
-   choose the state of interest, and load in the data using the code from the very top of this document (but replace `ca` twice with the two-digit letters for your state)


# Independent Practice

## 1. Tree Equity vs. Poverty

> Is there a relationship between poverty (`pctpov`) and tree equity score (`tes`)?


```{r}
#| echo: true
#| eval: false

ca_tes |>
  ___(___(x = ___, y = ___)) +
  ___(color = "green") 
```



## 2. Distribution of tree equity scores

> What is the distribution of tree equity scores (`tes`)?


```{r}
#| echo: true
#| eval: false

___ |>
  ggplot(aes(___)) +
  ___(fill = "red")
```



## 3. Distribution of tree equity scores

> What is the distribution of tree equity scores (`tes`) created by a boxplot?


```{r}
#| echo: true
#| eval: false

___ |>
  ggplot(aes(___)) +
  ___(fill = "violet")
```



## 4. Tree equity vs. linguistic isolation

> Is there a relationship between linguistic isolation(`lingnorm`) and `tes` in an area. 

  - Use ggplot to plot the relationship between `tes` and `lingnorm`
  - Choose the best plot for the variable types
  - Create fitting labels for your axes and plot title


## 5. Wrangle + plot

Recreate the plot above, but summarize the information **per `county`** before plotting. That is, `Los Angeles County` should now only be represented by one dot. Summarize both `lingnorm` and `tes` using the `mean()` function.


## 6. Wrangle + plot + label

Recreate the plot above, but now add labels to the plot to indicate the county of each dot.  Use `geom_label()` or `geom_text()`. Try both! You'll need to use `label = county`...  do you need `aes()`? 



## 7. Wrangle + plot + label

There is an R package called **ggrepel**.  It includes a geom called `geom_label_repel`. Try it!  (To use a function from a package, you can write: `ggrepel::geom_label_repel()`.)

**Hint:** If R can't find the **ggrepel** package, you may need to type `install.packages("ggrepel")` in the Console down below.


